Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-10-28 algorithm.tex
Fixed references

\subsection{Algorithm Parameters and Scalability} The parameters that
completely determine the algorithm are the configurations budget each
era $K$ and the threshold factor $\alpha$. From a functional point of
view, $K$ is somewhat similar to the population size of a genetic
approach, determining which configuration should be evaluated in the
in the subsequent step of the algorithm. However, as seen in the
previous subsection, the way this $K$ configurations are distributed
across the parameters space is completely different.
Authors of \cite{zitzler_ec00} found that
increasing $k$ not necessarily has a considerable impact on the
performance of the algorithm, and this holds also in our algorithm.
But while in the genetic approach $K$ must be sufficiently large
in order to guarantee diversity of the Pareto set configurations, in
our algorithm the diversity is not impacted by $K$ since, by
construction, it is guaranteed by the way the configuration space is
partitioned, independently from number of configurations evaluated at
each iteration.

On the other side, parameter $\alpha$ determines when a region is considered of high
innovation. High values of $\alpha$ mean that a region will hardly be
elected as high-innovation, yielding less splitting operations. This
slows down zooming in on portions of configuration space and make the
distribution of the simulated configuration more uniform, as there are
fewer small regions with a high test configuration concentration. The
opposite happens when small values of $\alpha$ are used. In this
introductory work we are not explicitly focused on investigating how
these parameters could be fine tuned, nevertheless
we conducted some preliminary tests to confirm the relative stability
of $PS$ with different choices and to select some ``reasonable'' average
values to perform the design space exploration (see for example Figure
TODO in Section \ref{sec:ee})
\comment{TODO This is confirmed by results of \secR{parameter_impact}
}).

>>>>

\subsection{Algorithm Parameters and Scalability} The parameters that
completely determine the algorithm are the configurations budget each
era $K$ and the threshold factor $\alpha$. From a functional point of
view, $K$ is somewhat similar to the population size of a genetic
approach, determining which configuration should be evaluated in the
subsequent step of the algorithm. However, as seen in the
previous subsection, the way this $K$ configurations are distributed
across the parameters space is completely different.
Authors of \cite{zitzler_ec00} found that
increasing $K$ not necessarily has a considerable impact on the
performance of the algorithm, and this holds also in our algorithm (see for example
Figure~\ref{fig:kimpact} in Subsection~\ref{sec:comparison}).
But while in the genetic approach $K$ must be sufficiently large
in order to guarantee diversity of the Pareto set configurations, in
our algorithm the diversity is not impacted by $K$ since, by
construction, it is guaranteed by the way the configuration space is
partitioned, independently from number of configurations evaluated at
each iteration.

On the other side, parameter $\alpha$ determines when a region is considered of high
innovation. High values of $\alpha$ mean that a region will hardly be
elected as high-innovation, yielding less splitting operations. This
slows down zooming in on portions of configuration space and make the
distribution of the simulated configuration more uniform, as there are
fewer small regions with a high test configuration concentration. The
opposite happens when small values of $\alpha$ are used. In this
introductory work we are not explicitly focused on investigating how
these parameters could be fine tuned, nevertheless
we conducted some preliminary tests to confirm the relative stability
of $PS$ with different choices and to select some ``reasonable'' average
values to perform the design space exploration, as it will discussed
in~\secR{ee}.

### 2014-10-28 experiments.tex
Fixed references

\subsection{Comparison Setup}

@@

\comment{TODO controllare inglese}
Even if an analogy between the concepts of ``era'' and ``generation'' 
can be considered, a fixed amount of total different simulations has been
considered for each of the tests. The purpose is to make the
comparison fair from a algorithm performance perspective, since a
fixed budget of simulations is directly related to the amount of time
required to execute the exploration; on the other side, simply
referring to number of generations/era could be misleading when a
significative number of configuration remain unaltered in the current
population (e.g. because of a local optimum).

The set of default algorithm tuning parameters used for both the
genetic and PS approaches to perform the simulations are shown in
Table~\ref{tab:defaults}. Given the similarity, both in terms of
meaning and functionality, of the parameter $K$ with the population size
of the genetic approach, they both have been set to the same values. As discussed
when introducing the main PS algorithm concepts, an accurate
investigation of algorithm tuning is beyond the scope of this
introductory work. Nevertheless, a set of preliminary tests with
different values of $K$ have been conducted. As shown in
Figure~\ref{fig:kimpact}, which refer to the alloca\_test application
and a fixed amount of 500 configurations, initially increasing values
of $K$ seem to improve both pareto (left picture) and objective
distribution (right picture), but after some threshold the $K$ does
not seem to provide any benefit, even returning to slightly worse
results for the last value considered (100). Similar trends have been
found for other benchmarks in the set considered (which are
not included for space reasons), showing a value of $K$ around 30 as
reasonable sweet spot for the objectives. A complete resource with the whole 
results and data obtained in the tests performed can be found
at~\cite{ps_results}.

>>>>

\subsection{Comparison Setup}
\secL{comparison}

@@

\comment{TODO controllare inglese}
Even if an analogy between the concepts of ``era'' and ``generation'' 
can be considered, a fixed amount of total different simulations has been
considered for each of the tests. The purpose is to make the
comparison fair from a algorithm performance perspective, since a
fixed budget of simulations is directly related to the amount of time
required to execute the exploration; on the other side, just simply
referring to a given number of generations/era could be misleading when a
significative number of configuration remain unaltered in the current
population (e.g. because of a local optimum).

The set of default algorithm tuning parameters used for both the
genetic and PS approaches to perform the simulations are shown in
Table~\ref{tab:defaults}. Given the similarity, both in terms of
meaning and functionality, of the parameter $K$ with the population size
of the genetic approach, they both have been set to the same values. As discussed
when introducing the main PS algorithm concepts, an accurate
investigation of algorithm tuning is beyond the scope of this
introductory work. Nevertheless, a set of preliminary tests with
different values of $K$ have been conducted. As shown in
Figure~\ref{fig:kimpact}, which refer to the alloca\_test application
and a fixed amount of 500 configurations, initially increasing values
of $K$ seem to improve both pareto (left picture) and objective
distribution (right picture), but after some threshold the $K$ does
not seem to provide any benefit, returning to slightly worse
results for the last value considered (100). Similar trends have been
found for other benchmarks in the set considered (which are
not included for space reasons), showing a value of $K$ around 30 as
reasonable sweet spot for the objectives. A complete resource with the whole 
results and data obtained in the tests performed can be found
at~\cite{ps_results}.
